---
title: "CA4bVdVeen"
author: "Ties van der Veen"
date: "22-9-2019"
output: pdf_document
---

## Rebecca Costello - Ties van der Veen - Niels van Opstal

# I. Introduction to the assignment

## Packages

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(foreign)
library(tidyverse)
library(ggdag)
library(dplyr)
library(tinytex)
library(jtools)
library(huxtable)
library(summarytools)
library(ggstance)
library(pwr)
library(knitr)
library(lemon)
library(AER)
library(rdrobust)
library(lubridate)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
knit_print.data.frame <- lemon_print
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
st_options(plain.ascii = FALSE, style = "rmarkdown")
st_css()
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
theUrl_ca4b <- "https://surfdrive.surf.nl/files/index.php/s/JAo4jnvf042xQXt/download"
fertility <- read.dta (file = theUrl_ca4b)
```

# II. What to submit

## 1. Mechanism

Y(0,i) = relatively lower levels of fertility, standard number of abortions and quicker re-entry into the labor force
Y(1,i) = increased fertility, reduction in abortions and slower re-entry into the labor force

Child benefits after a birth raises fertility because it raises expendable income, which means the parents need to work less hours to satisfy the same (monetary) needs. It also lowers the bar to sustain a child's upbringing, making it more likely that the parents decide in favour of getting more children.

## 2. Descriptives

(a)

```{r}
abspermonth <- fertility %>% group_by(month(fertility$abortions)) %>% tally()
mean(abspermonth$n)

```

Probably not right (as the paper says 6,725) but no idea how to fix it.

(b)

```{r}
conpermonth <- fertility %>% group_by(month(fertility$conceptions)) %>% tally()
mean(conpermonth$n)
```

B relies on A, no idea what I'm doing wrong with A so can't do B. Wanted to get the average conceptions, multiply average abortions by 1.14 (as it says in the question) and then compare percentages.

(c)

```{r}
fertility$date1 <- fertility$calendar_year+(fertility$calendar_month/12)
fertility$date1 <- fertility$date1-(1/12)
fertility$date1 <- as.yearmon(fertility$date1)
```

```{r}
fertility$date1 <- as.Date(fertility$date1)
min(fertility$date1, na.rm=TRUE)
max(fertility$date1, na.rm=TRUE)
```

(d)

```{r}
plot(fertility$date1, fertility$abortions,
     xlab="Date", ylab="Average number of abortions",
     ylim = c(0, 10000))
abline(lm(fertility$abortions ~ fertility$date1))
```

## 3. Graphical evidence

(a)


```{r}
with(fertility, rdplot(y=conceptions, p=1, x=month_of_conception[0:69], c=40, nbins = 35,
                     title="Discontinuity in fertility",
                     x.label="days relative to July 1, 2007",
                     y.label="number of conceptions"))
```

Don't understand why I'm getting this error. Think the rest is correct.

(b)

Can't do this without the graph

(c)

```{r}
with(fertility, rdplot(y=abortions, p=1, x=month_of_birth[0:69], c=40, nbins = 35,
                     title="Discontinuity in abortions",
                     x.label="days relative to July 1, 2007",
                     y.label="number of abortions"))
```

Don't understand the error.

## 4. Test of treatment effect

```{r}
reg4b1 <- lm(ln_conceptions ~ treatment_dummy, data=fertility)
summ(reg4b1)
```

```{r}
reg4b2 <- lm(ln_abortions ~ treatment_dummy, data=fertility)
summ(reg4b2)
```

Naïve regressions for both outcome variables.

Don't know how to do the rest without running into the same errors.




